A Horrocks-type theorem for even orthogonal $K_2$

### 2020-04-15 dl-paper.tex
Minor fix
Some further fixes

Recall that to every irreducible root system $\Phi$ and a commutative ring $R$ one can associate two groups: the {\it simply-connected Chevalley group} $\GG(\Phi, R)$ (see e.\,g.~\cite[\S~3]{St71} or~\cite{VP}) and the {\it Steinberg group} $\St(\Phi, R)$ (see~\cref{sec:Steinberg-intro} for the definition). There is a well-defined homomorphism $\pi \colon \St(\Phi, R) \to \GG(\Phi, R)$ sending each generator $x_\alpha(\xi)$ to the elementary root unipotent $t_\alpha(\xi)$. The cokernel and the kernel of this homomorphism (denoted $\K_1(\Phi, R)$ and $\K_2(\Phi, R)$, respectively) are, by definition, the values of the so-called $\mathrm{K}_1$ and $\mathrm{K_2}$-functors modeled on the Chevalley group $\GG(\Phi, -)$, see~\cite{St78}.

@@

\section{Preliminaries}
If $x$ and $y$ are elements of a group $G$ then $[x, y]$ denotes the left-normed commutator $xyx^{-1}y^{-1}$.
We denote by $x^y$ (resp. ${}^y\!x$) the element $y^{-1}xy$ (resp. $yxy^{-1}$). 
Throughout the paper we make use of the following commutator identities:
\begin{align}
 [x, yz]^y =& [y^{-1}, x] \cdot [x, z], \label{rel43} \\ 
 \label{eq:H1ii} [xy, z] =& {}^x[y, z] \cdot [x,z], \\
 \label{eq:H1ii-2}[x, yz] =& [x, y] \cdot {}^{y}\![x, z], \\
 [[x, y], z] =& \left([x^{-1}, [ y^{-1}, z]] ^ {y^{-1}} \cdot [y, [ z^{-1}, x^{-1}]] ^ {z^{-1}} \right)^{x^{-1}}. \label{HW-variant} \end{align}
Recall also from~\cite[Lemma~3.1.1]{RS76} that
\begin{equation} \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz]. \end{equation}

@@

For the rest of this section let us fix the following notation.
Let $A$ be an arbitrary commutative local ring with maximal ideal $M$ and residue field $k$.
Denote by $B$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$ and
by $I$ the ideal $M[X, X^{-1}]$ of $B$ (it is clear that $I$ is also an ideal of $R$).

@@

\begin{lemma} \label{Crels} Let $\Phi$ be a simply-laced root system.
The elements $c_\alpha(s, t)$ satisfy the following relations for all $s\in I,\ t\in J,\ \xi\in R$:
 \begin{enumerate} [label=\normalfont(\arabic*)]
 \item \label{C1} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta}\cdot s^2t\xi)$ if $\alpha+\beta \in \Phi$;
 \item \label{C2} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(st\xi + s^2t^2\xi) \cdot x_{\alpha-\beta}(N_{-\alpha, \beta}\cdot st^2\xi)$ if $\alpha-\beta \in \Phi$;  
 \item \label{C3} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = 1$ if $\alpha \perp \beta$;  
 \item \label{C4} If $\alpha+\beta\in\Phi$ then
  \begin{equation} \label{eq:C4} c_{\alpha+\beta}(s, t\xi) = [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2),\end{equation}
  where $\epsilon = N_{\alpha,\beta}$.
 \end{enumerate}
\end{lemma}
\begin{proof}
The first two assertions can be verified by direct calculation:
\begin{align*} \nonumber [c_{\beta}(s, t),\ x_{\alpha}(\xi)] = [x_{\beta}(s),\ x_{-\beta}(t)] \cdot [x_{-\beta}(t),\ x_{\beta}(s)\cdot x_{\alpha + \beta}(N_{\alpha, \beta} \cdot s\xi)] &\text{ by~\eqref{Cdef}, \eqref{Chevalley-CCF1}, \eqref{Chevalley-CCF2}} \\
 = {}^{x_\beta(s)}\![x_{-\beta}(t),\ x_{\alpha+\beta}(N_{\alpha, \beta} \cdot s\xi)] &\text{ by~\eqref{eq:H1ii-2}} \\ = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta} \cdot s^2t\xi) &\text{ by~\eqref{Chevalley-CCF1};} \end{align*}
\begin{align*} \nonumber
\begin{split}[[x_\beta(s),\ x_{-\beta}(t)],\ x_{\alpha}(\xi)] = \bigl([x_\beta(-s),\ [x_{-\beta}(-t),\ x_{\alpha}(\xi)]] ^ {x_{-\beta}(-t)} \cdot \hspace{50pt} \\ \cdot [x_{-\beta}(t),\ [ x_{\alpha}(-\xi),\ x_\beta(-s)]] ^ {x_{\alpha}(-\xi)}\bigr)^{x_\beta(-s)} & \text{ by~\eqref{HW-variant}}\end{split} \\
= x_{\alpha}(-N_{\beta, \alpha-\beta} N_{\alpha,-\beta} \cdot s t \xi) ^ {x_{-\beta}(-t) \cdot x_\beta(-s)} = x_{\alpha}(s t \xi) ^ {x_{-\beta}(-t) \cdot x_\beta(-s)} & \text{ by~\eqref{Chevalley-CCF1},\eqref{Chevalley-CCF2}}\\
= x_{\alpha}(st\xi) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot st^2\xi)^{x_\beta(-s)} = x_{\alpha}(st\xi + s^2t^2 \xi) \cdot x_{\alpha-\beta}(N_{-\alpha,\beta}\cdot st^2\xi) & \text{ by~\eqref{Chevalley-CCF1}.} \end{align*}
The assertion~\ref{C3} is obvious. Finally, last assertion can be obtained as follows:
\begin{align*}
[x_{\alpha+\beta}(s),\ x_{-\alpha-\beta}(t\xi)] = [x_{\alpha+\beta}(s),\ [x_{-\alpha}(-\epsilon t),\ x_{-\beta}(\xi)]]&\ \text{by~\eqref{Chevalley-CCF1}}\\
\begin{split}
    =\bigl( [[x_{\alpha+\beta}(-s),\ x_{-\alpha}(\epsilon t)],\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s)} \cdot \hspace{70pt} \\
        \cdot [[ x_{-\beta}(-\xi),\ x_{\alpha+\beta}(s)],\ x_{-\alpha}(\epsilon t)]^{x_{-\beta}(-\xi)}\bigr)^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{HW-variant}}
\end{split}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t)]^{x_{-\beta}(-\xi)x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t) x_{-\alpha-\beta}(t\xi)]^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1},\eqref{Chevalley-CCF2}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{-\alpha}(-\epsilon t),\ x_{\alpha}(\epsilon s\xi)] \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha-\beta}(t\xi)] &\ \text{by~\eqref{rel43}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2)  &\ \text{by~\eqref{Chevalley-CCF1}}. \qedhere
\end{align*}

@@

Throughout this section we use the following notations and conventions:
\begin{itemize}
 \item $A$ denotes an arbitarary commutative ring and $M$ is an ideal of $A$. Starting from subsection~\ref{sec:Pa0-basic} we also assume that $A$ is local and $M$ is the maximal ideal of $A$.
 \item We denote by $R$ the Laurent polynomial ring $A[X, X^{-1}]$. We set $t = X^{-1}$, so that $R = A[X, X^{-1}] = A[t, t^{-1}]$.
 \item $B$ denotes the subring $A[t] + M[t^{-1}] = A[X^{-1}] + M[X]$ of $R$.
 \item $I$ denotes the ideal $M[X, X^{-1}]$ of $R$. Notice that $I$ is also an ideal of $B$.
 \end{itemize}
\subsection{Presentation of Steinberg groups by homogeneous generators}
\label{sec:presentation}
We consider $R = A[t, t\inv]$ as a $\mathbb{Z}$-graded ring in which $t$ has degree $1$. This grading induces the grading on the subring $B \subseteq R$. For an integer $d$ we denote by $R_d$ (resp. $B_d$) the degree $d$ part of the ring $R$ (resp. $B$). Obviously, $B_d=M \cdot t^d$ for $d<0$, and $B_d=A \cdot t^d$ for $d\geq0$. With this notation, $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ as an $A$-module.

@@

 Now let $A$ be arbitrary commutative ring and $\Phi = \rD_\ell$, $\ell \geq 7$. If $g \in \St(\Phi, A[X])$ is such that its image in $\St(\Phi, A[X, X\inv])$ is trivial, then so is its image in all localizations $\St(\Phi, A_M[X, X\inv])$, where $M$ ranges over the maximal ideals of $A$.
 We denote by $\lambda_{M}$ (resp. $\lambda_{M, -}$) the localization homomorphism $A[X] \to A_M[X]$ (resp. $A[X\inv] \to A_M[X\inv]$) and by $\lambda_M^*$, $\lambda_{M,-}^*$ the corresponding homomorphism of Steinberg groups.
 By the just established local case of the theorem the images $\lambda^*_M(g)$ in all $\St(\Phi, A_M[X])$ are also trivial. Now by the local-global principle \cite[Theorem~2]{LS17} the element $g$ is trivial as well.

>>>>

Recall that to every irreducible root system $\Phi$ and a commutative ring $R$ one can associate two groups: the {\it simply-connected Chevalley group} $\GG(\Phi, R)$ (see e.\,g.~\cite[\S~3]{St71} or~\cite{VP}) and the {\it Steinberg group} $\St(\Phi, R)$ (see~\cref{sec:Steinberg-intro} for the definition). There is a well-defined homomorphism $\pi \colon \St(\Phi, R) \to \GG(\Phi, R)$ sending each generator $x_\alpha(\xi)$ to the elementary root unipotent $t_\alpha(\xi)$. The cokernel and the kernel of this homomorphism are denoted $\K_1(\Phi, R)$ and $\K_2(\Phi, R)$, respectively. The latter groups are functorial in $R$. The functors $\K_1(\Phi, -)$ and $\K_2(\Phi, -)$ are called {\it the $\mathrm{K}_1$ and $\mathrm{K_2}$-functors modeled on the Chevalley group $\GG(\Phi, -)$}, see~\cite{St78}.

@@

\section{Preliminaries}
If $x$ and $y$ are elements of a group $G$ then $[x, y]$ denotes the left-normed commutator $xyx^{-1}y^{-1}$.
We denote by $x^y$ (resp. ${}^y\!x$) the element $y^{-1}xy$ (resp. $yxy^{-1}$). 
Throughout the paper we make use of the following commutator identities:
\begin{align}
 [x, yz]^y =& [y^{-1}, x] \cdot [x, z], \label{rel43} \\ 
 \label{eq:H1ii} [xy, z] =& {}^x[y, z] \cdot [x,z], \\
 \label{eq:H1ii-2}[x, yz] =& [x, y] \cdot {}^{y}\![x, z], \\
 [[x, y], z] =& {}^{x}\!\left({}^{y}[x^{-1}, [ y^{-1}, z]] \cdot {}^z[y, [ z^{-1}, x^{-1}]] \right). \label{HW-variant} \end{align}
Recall also from~\cite[Lemma~3.1.1]{RS76} that
\begin{equation} \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz]. \end{equation}

@@

For the rest of this section let us fix the following notation.
Let $A$ be an arbitrary commutative local ring with maximal ideal $M$ and residue field $k$.
Denote by $B$ the subring $A[X^{-1}] + M[X]$ of the ring $R = A[X, X^{-1}]$ and
by $I$ the ideal $M[X, X^{-1}]$ of $R$ (it is clear that $I$ is also an ideal of $B$).

@@

\begin{lemma} \label{Crels} Let $\Phi$ be a simply-laced root system.
The elements $c_\alpha(s, t)$ satisfy the following relations for all $s\in I,\ t\in J,\ \xi\in R$:
 \begin{enumerate} [label=\normalfont(\arabic*)]
 \item \label{C1} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta}\cdot s^2t\xi)$ if $\alpha+\beta \in \Phi$;
 \item \label{C2} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = x_{\alpha}(st\xi + s^2t^2\xi) \cdot x_{\alpha-\beta}(N_{-\alpha, \beta}\cdot st^2\xi)$ if $\alpha-\beta \in \Phi$;  
 \item \label{C3} $[c_\beta(s, t),\ x_{\alpha}(\xi)] = 1$ if $\alpha \perp \beta$;  
 \item \label{C4} If $\alpha+\beta\in\Phi$ then
  \begin{equation} \label{eq:C4} c_{\alpha+\beta}(s, t\xi) = [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2),\end{equation}
  where $\epsilon = N_{\alpha,\beta}$.
 \end{enumerate}
\end{lemma}
\begin{proof}
The first two assertions can be verified by direct calculation:
\begin{align*} \nonumber [c_{\beta}(s, t),\ x_{\alpha}(\xi)] = [x_{\beta}(s),\ x_{-\beta}(t)] \cdot [x_{-\beta}(t),\ x_{\beta}(s)\cdot x_{\alpha + \beta}(N_{\alpha, \beta} \cdot s\xi)] &\text{ by~\eqref{Cdef}, \eqref{Chevalley-CCF1}, \eqref{Chevalley-CCF2}} \\
 = {}^{x_\beta(s)}\![x_{-\beta}(t),\ x_{\alpha+\beta}(N_{\alpha, \beta} \cdot s\xi)] &\text{ by~\eqref{eq:H1ii-2}} \\ = x_{\alpha}(- st\xi) \cdot x_{\alpha+\beta}(N_{\alpha,\beta} \cdot s^2t\xi) &\text{ by~\eqref{Chevalley-CCF1};} \end{align*}
\begin{align*} \nonumber
\begin{split}[[x_\beta(s),\ x_{-\beta}(t)],\ x_{\alpha}(\xi)] = {}^{x_\beta(s)}\!\bigl( {}^{x_{-\beta}(t)}[x_\beta(-s),\ [x_{-\beta}(-t),\ x_{\alpha}(\xi)]] \cdot \hspace{50pt} \\ \cdot {}^{x_{\alpha}(\xi)}[x_{-\beta}(t),\ [ x_{\alpha}(-\xi),\ x_\beta(-s)]] \bigr)^{} & \text{ by~\eqref{HW-variant}}\end{split} \\
= {}^{x_\beta(s) \cdot x_{-\beta}(t)}x_{\alpha}(-N_{\beta, \alpha-\beta} N_{\alpha,-\beta} \cdot s t \xi) = {}^{ x_\beta(s) \cdot x_{-\beta}(t)}x_{\alpha}(s t \xi) & \text{ by~\eqref{Chevalley-CCF1},\eqref{Chevalley-CCF2}}\\
= x_{\alpha}(st\xi) \cdot x_{\alpha-\beta}(-N_{\alpha,-\beta}\cdot st^2\xi)^{x_\beta(-s)} = x_{\alpha}(st\xi + s^2t^2 \xi) \cdot x_{\alpha-\beta}(N_{-\alpha,\beta}\cdot st^2\xi) & \text{ by~\eqref{Chevalley-CCF1}.} \end{align*}
The assertion~\ref{C3} is obvious. Finally, the last assertion can be obtained as follows:
\begin{align*}
[x_{\alpha+\beta}(s),\ x_{-\alpha-\beta}(t\xi)] = [x_{\alpha+\beta}(s),\ [x_{-\alpha}(-\epsilon t),\ x_{-\beta}(\xi)]]&\ \text{by~\eqref{Chevalley-CCF1}}\\
\begin{split}
    ={}^{x_{-\alpha}(-\epsilon t)}\!\bigl( {}^{x_{\alpha+\beta}(s)}[[x_{\alpha+\beta}(-s),\ x_{-\alpha}(\epsilon t)],\ x_{-\beta}(\xi)]  \cdot \hspace{70pt} \\
        \cdot {}^{x_{-\beta}(\xi)}[[ x_{-\beta}(-\xi),\ x_{\alpha+\beta}(s)],\ x_{-\alpha}(\epsilon t)]\bigr) &\ \text{by~\eqref{HW-variant}}
\end{split}\\
= \left({}^{x_{\alpha+\beta}(s)}[x_{\beta}(st),\ x_{-\beta}(\xi)] \cdot {}^{x_{-\beta}(\xi)}[x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t)]\right)^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1}}\\
= \left({}^ {x_{\alpha+\beta}(s)}[x_{\beta}(st),\ x_{-\beta}(\xi)] \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha}(\epsilon t) \cdot x_{-\alpha-\beta}(t\xi)]\right)^{x_{-\alpha}(\epsilon t)} &\ \text{by~\eqref{Chevalley-CCF1},\eqref{Chevalley-CCF2}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot [x_{-\alpha}(-\epsilon t),\ x_{\alpha}(\epsilon s\xi)] \cdot [x_{\alpha}(\epsilon s\xi),\ x_{-\alpha-\beta}(t\xi)] &\ \text{by~\eqref{rel43}}\\
= [x_{\beta}(st),\ x_{-\beta}(\xi)] ^ {x_{\alpha+\beta}(-s) x_{-\alpha}(\epsilon t)} \cdot c_{\alpha}(\epsilon s\xi, -\epsilon t)^{-1} \cdot x_{-\beta}(-st\xi^2)  &\ \text{by~\eqref{Chevalley-CCF1}}. \qedhere
\end{align*}

@@

Throughout this section we use the following notations and conventions:
\begin{itemize}
 \item $A$ denotes an arbitarary commutative ring and $M$ is an ideal of $A$. Starting from subsection~\ref{sec:Pa0-basic} we also assume that $A$ is local and $M$ is the maximal ideal of $A$.
 \item We denote by $R$ the Laurent polynomial ring $A[X, X^{-1}]$. We set $t = X^{-1}$, so that $R = A[X, X^{-1}] = A[t, t^{-1}]$.
 \item $B$ denotes the subring $A[t] + M[t^{-1}] = A[X^{-1}] + M[X]$ of $R$.
 \item $I$ denotes the ideal $M[X, X^{-1}]$ of $R$ (it is clear that $I$ is also an ideal of $B$).
 \end{itemize}
\subsection{Presentation of Steinberg groups by homogeneous generators}
\label{sec:presentation}
We consider $R = A[t, t\inv]$ as a $\mathbb{Z}$-graded ring in which $t$ has degree $1$. This grading induces the grading on the subring $B \subseteq R$. For an integer $d$ we denote by $R_d$ (resp. $B_d$) the degree $d$ part of the ring $R$ (resp. $B$). Obviously, $B_d=M \cdot t^d$ for $d<0$, and $B_d=A \cdot t^d$ for $d\geq0$. With this notation, $B$ decomposes as $\oplus_{d\in\mathbb Z}B_d$ as an $A$-module.

@@

 Now let $A$ be arbitrary commutative ring and $\Phi = \rD_\ell$, $\ell \geq 7$. If $g \in \St(\Phi, A[X])$ is such that its image in $\St(\Phi, A[X, X\inv])$ is trivial, then so is its image in all localizations $\St(\Phi, A_M[X, X\inv])$, where $M$ ranges over the maximal ideals of $A$.
 We denote by $\lambda_{M}$ (resp. $\lambda_{M, -}$) the localization homomorphism $A[X] \to A_M[X]$ (resp. $A[X\inv] \to A_M[X\inv]$) and by $\lambda_M^*$, $\lambda_{M,-}^*$ the corresponding homomorphisms of Steinberg groups.
 By the just established local case of the theorem the images $\lambda^*_M(g)$ in all $\St(\Phi, A_M[X])$ are also trivial. Now by the local-global principle \cite[Theorem~2]{LS17} the element $g$ is trivial as well.
